\Large Feature and order manipulations in a free recall task affect memory for current and future lists

### 2023-11-20 paper/main.tex
moved most stats to tables

%! suppress = LineBreak
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{geometry}
\usepackage[sort&compress]{natbib}
\usepackage{pxfonts}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{lineno}
\usepackage{amsopn}
\usepackage{rotating}

@@

\newcommand{\dynamicsRandom}{S1}
\newcommand{\dynamicsAdaptive}{S2}
\newcommand{\accuracyByList}{S3}
\newcommand{\clusterCorrs}{S4}
\newcommand{\fingerprintsRandom}{S5}
\newcommand{\fingerprintsAdaptive}{S6}
\newcommand{\recallInit}{S7}
\newcommand{\fingerprintTrajectoryRandom}{S8}
\newcommand{\featureCorrs}{S9}

@@

To obtain a clean preliminary estimate of the consequences on memory of
randomly varying the font colors and locations of presented words (versus
holding the font color fixed at black, and the words' locations fixed at the
center of the screen), we compared participants' performance on the
\textit{feature-rich} and \textit{reduced} experimental conditions (see
\textit{Random conditions}, Fig.~\dynamicsRandom). In the feature-rich
condition, the words' colors and locations varied randomly, and in the reduced
condition, words were always presented in black, at the center of the display.
Aggregating across all lists for each participant, we found no difference in
recall accuracy (i.e., the proportions of words successfully recalled) for
feature-rich versus reduced lists ($t(126) = -0.290,~p = 0.772$, Cohen's~$d =
-0.051$, bootstrap-estimated 95\% confidence interval (CI)~$= [-2.387,~
1.768]$). However, participants in the feature-rich condition clustered their
recalls substantially more along every dimension we examined (temporal
clustering: $t(126) = 10.632,~p < 0.001,~d = 1.882,~\mathrm{CI} = [7.786,~
14.386]$; semantic category clustering: $t(126) = 10.148,~p < 0.001,~d =
1.796,~ \mathrm{CI} = [7.324,~13.778]$; size clustering: $t(126) = 12.033,~p <
0.001,~ d = 2.129,~\mathrm{CI} = [9.030,~15.918]$; word length clustering:
$t(126) = 10.720,~p < 0.001,~d = 1.897,~\mathrm{CI} = [7.442,~15.174]$; first
letter clustering: $t(126) = 6.679,~p < 0.001,~d = 1.182,~\mathrm{CI} =
[4.490,~ 9.611]$; see \textit{Permutation-corrected feature clustering scores}
for more information about how we quantified each participant's clustering
tendencies.) Taken together, these comparisons suggest that adding new features
changes how participants organize their memories of studied words, even when
those new features are independent of the words themselves and vary randomly
across words. We found no evidence that those additional uninformative features
were distracting (in terms of their impact on memory performance), but they did
affect participants' recall dynamics (measured via their clustering scores).

@@

We next asked whether adding these incidental visual features to later lists
(after the participants had already studied impoverished lists), or removing
the visual features from later lists (after the participants had already
studied visually diverse lists) might affect memory performance. In other
words, we sought to test for potential effects of changing the ``richness'' of
participants' experiences over time. All participants studied and recalled a
total of 16 lists; we defined \textit{early} lists as the first eight lists and
\textit{late} lists as the last eight lists each participant encountered. To
help interpret our results, we compared participants' memories on early versus
late lists in the above feature-rich and reduced conditions. Participants in
both conditions remembered more words on early versus late lists (feature-rich:
$t(66) = 4.553,~p < 0.001,~d = 0.233,~\mathrm{CI} = [2.427,~7.262]$; reduced:
$t(60) = 2.434,~p = 0.018,~d = 0.134,~\mathrm{CI} = [0.493,~4.910]$).
Participants in the feature-rich (but not reduced) conditions exhibited more
temporal clustering on early versus late lists (feature-rich: $t(66) = 2.268,~p
= 0.027,~d = 0.181,~\mathrm{CI} = [0.437,~4.425]$; reduced: $t(60) = 0.986,~p =
0.328,~d = 0.061,~\mathrm{CI} = [-0.897,~3.348]$). And participants in both
conditions tended to exhibit more semantic clustering on early versus late
lists (feature-rich, category: $t(66) = 3.684,~p < 0.001,~d = 0.220,~
\mathrm{CI} = [1.733,~5.732]$; feature-rich, size: $t(66) = 1.629,~p = 0.108,~d
= 0.100,~\mathrm{CI} = [-0.207,~3.905] $; reduced, category: $t(60) = 2.755,~p
= 0.008,~d = 0.177,~\mathrm{CI} = [0.761,~5.189]$; reduced, size: $t(60) =
3.081,~p = 0.003,~d = 0.201,~\mathrm{CI} = [1.210,~5.326]$). Participants in
the reduced (but not feature-rich) conditions tended to exhibit more
lexicographic clustering on early versus late lists (feature-rich, word length:
$t(66) = -0.100,~p = 0.921,~d = -0.010,~\mathrm{CI} = [-2.217,~1.899]$; feature
rich, first letter: $t(66) = 0.412,~p = 0.681,~d = 0.045,~\mathrm{CI} =
[-1.645,~2.461]$; reduced, word length: $t(60) = 3.762,~p < 0.001,~d = 0.261,~
\mathrm{CI} = [1.604,~6.821]$; reduced, first letter: $t(60) = 1.721,~p =
0.090,~d = 0.175,~\mathrm{CI} = [-0.138,~4.098]$). Taken together, these
comparisons suggest that even when the presence or absence of incidental visual
features is stable across lists, participants still exhibit some differences in
their performance and memory organization tendencies for early versus late
lists.

With these differences in mind, we next compared participants' memories on
early versus late lists for two additional experimental conditions (see
\textit{Random conditions}, Fig.~\dynamicsRandom). In a \textit{reduced
(early)} condition, we held the visual features constant on early lists, but
allowed them to vary randomly on late lists. In a \textit{reduced (late)}
condition, we allowed the visual features to vary randomly on early lists, but
held them constant on late lists. Given our above findings that (a)
participants tended to exhibit stronger clustering effects on feature-rich
(versus reduced) lists, and (b) participants tended to remember more words and
exhibit stronger clustering effects on early (versus late) lists, we expected
these early versus late differences to be enhanced in the reduced (early)
condition and diminished in the reduced (late) condition. However, to our
surprise, participants in \textit{neither} condition exhibited reliable early-versus-late
differences in accuracy (reduced (early): $t(41) = 1.499, ~p =
0.141,~d = 0.098,~\mathrm{CI} = [-0.345,~3.579]$; reduced (late): $t(40) =
1.462,~p = 0.152,~d = 0.121,~\mathrm{CI} = [-0.376,~2.993]$), temporal
clustering (reduced (early): $t(41) = 0.857,~p = 0.396,~d = 0.068,~\mathrm{CI}
= [-1.012,~2.896]$; reduced (late): $t(40) = 1.244,~p = 0.221,~d = 0.128,~
\mathrm{CI} = [-0.894,~3.088]$), nor feature-based clustering (reduced (early),
category: $t(41) = 0.707,~p = 0.484,~d = 0.068,~\mathrm{CI} = [-1.314,~2.830]$;
reduced (early), size: $t(41) = 0.803,~p = 0.427,~d = 0.079,~\mathrm{CI} =
[-1.142,~2.953]$; reduced (early), length: $t(41) = 0.461,~p = 0.648,~d =
0.060,~\mathrm{CI} = [-1.545,~2.462]$; reduced (early), first letter: $t(41) =
0.781,~p = 0.439,~d = 0.101,~\mathrm{CI} = [-1.039,~2.881]$; reduced (late),
category: $t(40) = 0.101,~p = 0.920,~d = 0.009,~\mathrm{CI} = [-1.776,~2.307]$;
reduced (late), size: $t(40) = 0.555,~p = 0.582,~d = 0.058,~\mathrm{CI} =
[-1.444,~2.274]$; reduced (late), length: $t(40) = 1.482,~p = 0.146,~d =
0.126,~\mathrm{CI} = [-0.444,~3.743]$; reduced (late), first letter: $t(40) =
-0.143,~p = 0.887,~d = -0.017,~\mathrm{CI} = [-2.204,~1.830] $). We
hypothesized that adding or removing the variability in the visual features was
acting as a sort of ``event boundary'' between early and late
lists~\citep[e.g.,][]{ClewEtal19,RadvZack17,RadvCope06}. In prior work, we
(and others) have found that memories formed just after event boundaries can be
enhanced~\citep[e.g., due to less contextual interference between pre- and
post-boundary items;][]{MannEtal16, PettEtal16, GoldEtal17, FlorEtal17}.

We found that \textit{adding} incidental visual features on later lists that
had not been present on early lists (as in the reduced (early) condition)
served to enhance recall performance relative to conditions where all lists had
the same blends of features (accuracy for feature-rich vs.~reduced (early):
$t(107) = -2.230,~p = 0.028,~d = -0.439,~\mathrm{CI} = [-4.252,~-0.229]$;
reduced vs.~reduced (early): $t(101) = -2.045,~p = 0.043,~d = -0.410,~
\mathrm{CI} = [-3.826,~0.112]$; also see Fig.~\accuracyByList A). However,
\textit{subtracting} irrelevant visual features on later lists that
\textit{had} been present on early lists (as in the reduced (late) condition)
did not appear to impact recall performance (accuracy for feature-rich vs.~
reduced (late): $t(106) = -0.638,~p = 0.525,~d = -0.126,~\mathrm{CI} = [-2.720,~
1.362]$; reduced vs.~reduced (late): $t(100) = -0.407,~p = 0.685,~d =
-0.082,~\mathrm{CI} = [-2.477,~1.626]$). These comparisons suggest that recall
accuracy has a directional component: accuracy is affected differently by
removing features that had initially been present versus adding features
that had initially been absent. In contrast, we found that
participants exhibited more temporal and feature-based clustering when we added
incidental visual features to \textit{any} lists (comparisons of clustering on
feature-rich vs.~reduced lists are reported above; temporal clustering in
reduced vs.~reduced (early) and reduced vs.~reduced (late) conditions:
$t$s $\leq -9.885$,~$p$s $< 0.001$; feature-based clustering in reduced versus
reduced (early) and reduced versus reduced (late) conditions: $t$s $\leq
-4.555$,~$p$s $< 0.001$). Temporal and feature-based clustering were not
reliably different in the feature-rich, reduced (early), and reduced (late)
conditions (temporal clustering in feature-rich vs.~reduced (early) and
feature-rich vs.~reduced (late) conditions: $t$s $\geq -1.379$,~$p$s $\geq
0.171$; feature-based clustering in feature-rich vs.~reduced (early) and
feature-rich vs.~reduced (late) conditions: $|t|$s $\leq 1.441$,~$p$s $\geq
0.153$).

@@

@@

letter, color, and location). Participants in the category-ordered condition
showed an increase in memory performance on early lists (accuracy relative to
early feature-rich lists; $t(95) = 3.034,~p = 0.003,~d = 0.667,~\mathrm{CI} =
[1.048,~5.113]$). Participants in the color-ordered condition also showed a
trending increase in memory performance on early lists (again, relative to
early feature-rich lists: $t(96) = 1.850,~p = 0.067,~d = 0.402,~\mathrm{CI} =
[-0.010,~3.712]$; Fig.~\ref{fig:barplots}A). Participants' performances on
early lists in all of the other order manipulation conditions were
indistinguishable from performance on the early feature-rich lists ($|t|$s
$\leq 1.013,~p$s $\geq 0.314$). Participants in both of the semantically
ordered conditions exhibited stronger temporal clustering on early lists
(vs.~early feature-rich lists; category: $t(95) = 8.813,~p < 0.001,~d =
1.936,~\mathrm{CI} = [6.793,~11.751] $; size: $t(95) = 2.630,~p = 0.010,~d =
0.578,~\mathrm{CI} = [0.831,~4.866]$; Fig.~\ref{fig:barplots}B). Participants
in the length-ordered condition tended to exhibit \textit{less} temporal
clustering on early lists relative to early feature-rich lists ($t(95) =
-1.547,~p = 0.125,~d = -0.340,~\mathrm{CI} = [-3.693,~0.341]$), whereas
participants in the first letter-ordered condition exhibited stronger temporal
clustering on early lists ($t(95) = 2.858,~p = 0.005,~d = 0.628,~\mathrm{CI} =
[1.031,~4.886]$). Participants in the visually ordered conditions exhibited
more similar performance (accuracy) on early lists relative to early feature
rich lists (we found a trending enhancement for participants in the
color-ordered condition: $t(96) = 1.850,~p = 0.067,~d = 0.402,~\mathrm{CI} =
[-0.010,~3.712]$; location: $t(95) = 0.043,~p = 0.966,~d = 0.010,~\mathrm{CI} =
[-1.598,~1.729]$). Participants in the visually ordered conditions also showed
similar temporal clustering on early lists relative to early feature-rich
lists (color: $t(96) = -1.339,~p = 0.184,~d = -0.291,~\mathrm{CI} = [-3.238,~
0.394]$, we found a trending increase for participants in the location-ordered
condition: $t(95) = 1.705,~p = 0.092,~d = 0.374,~\mathrm{CI} = [-0.155,~
3.521]$). We also compared feature-based clustering on early lists across the
order manipulation and feature-rich conditions. Since these results were
similar across both semantic conditions (category and size), both lexicographic
conditions (length and first letter), and both visual conditions (color and
location), here we aggregate data from conditions that manipulated each of
these three feature groupings in our comparisons, to simplify the presentation.
On early lists, participants in the semantically ordered conditions exhibited
stronger semantic clustering relative to participants in the feature-rich
condition (category: $t(125) = 2.722,~p = 0.007,~d = 0.484,~\mathrm{CI} =
[0.827,~4.932]$; size: $t(125) = 3.866,~p < 0.001,~d = 0.687,~\mathrm{CI} =
[2.020,~5.983]$), but showed no reliable differences in lexicographic (length:
$t(125) = 0.521,~p = 0.603,~d = 0.093,~\mathrm{CI} = [-1.311,~2.333]$; first
letter: $t(125) = -0.842,~p = 0.401,~d = -0.150,~\mathrm{CI} = [-2.825,~
1.095]$) or visual (color: $t(125) = -0.650,~p = 0.517,~d = -0.116,~\mathrm{CI}
= [-2.680,~1.249]$; location: $t(125) = -0.251,~p = 0.802,~d = -0.045,
\mathrm{CI} = [-2.257,~1.524]$) clustering. Similarly, participants in the
lexicographically ordered conditions exhibited stronger (relative to feature
rich participants) lexicographic clustering (length: $t(125) = 3.682,~p <
0.001,~d = 0.655,~\mathrm{CI} = [1.890,~5.569]$; first letter: $t(125) = 5.134,
p < 0.001,~d = 0.912,~\mathrm{CI} = [3.251,~7.258]$) on early lists, but showed
no reliable differences in semantic (category: $t(125) = -1.040,~p = 0.301,~d =
-0.185,~\mathrm{CI} = [-3.095,~1.092]$; size: $t(125) = 0.006,~p = 0.995,~d =
0.001,~\mathrm{CI} = [-1.933,~1.952]$) or visual (color: $t(125) = 0.092,~p =
0.927,~d = 0.016,~\mathrm{CI} = [-1.834,~1.867]$; location: $t(125) = 0.407, p
= 0.685,~d = 0.072,~\mathrm{CI} = [-1.655,~2.463]$) clustering. And
participants in the visually ordered conditions exhibited stronger visual
clustering (again, relative to feature-rich participants, and on early lists;
color: $t(126) = 2.022,~p = 0.045,~d = 0.358,~\mathrm{CI} = [0.056,~3.965]$;
location: $t(126) = 4.390,~p < 0.001,~d = 0.777,~\mathrm{CI} = [2.730,~6.199]
$), but showed no reliable differences in semantic (category: $t(126) = 0.012,
p = 0.991,~d = 0.002,~\mathrm{CI} = [-1.988,~1.871]$; size: $t(126) = -0.104, p
= 0.917,~d = -0.018,~\mathrm{CI} = [-2.166,~1.847] $) or lexicographic (length:
$t(126) = 0.592,~p = 0.555,~d = 0.105,~\mathrm{CI} = [-1.361,~2.420]$; first
letter: $t(126) = 0.040,~p = 0.968,~d = 0.007,~\mathrm{CI} = [-1.791,~1.863]$)
clustering. Taken together, these order manipulation results suggest several
broad patterns (Figs.~\ref{fig:recall-dynamics}A,~\ref{fig:fingerprints}).
First, most of the order manipulations we carried out did \textit{not} reliably
affect overall recall performance. Second, most of the order manipulations
increased participants' tendencies to temporally cluster their recalls. Third,
all of the order manipulations enhanced participants' clustering of each
condition's target feature (i.e., semantic manipulations enhanced semantic
clustering, lexicographic manipulations enhanced lexicographic clustering, and
visual manipulations enhanced visual clustering; Fig.~\ref{fig:barplots}C)
while leaving clustering along other feature dimensions roughly unchanged
(i.e., semantic manipulations did not affect lexicographic or visual
clustering, and so on). Although it is not possible to fully separate feature-based
versus temporal clustering when considering sorted lists, we used a
permutation-based procedure to identify the degree of feature clustering over
and above what could be accounted for by temporal clustering alone (see
\textit{Factoring out the effects of temporal clustering}). When we carried out
this analysis (Fig.~\ref{fig:barplots}D), we found that participants exhibited
more semantic clustering on semantically sorted lists than on randomly ordered
lists, but the effects of the other order manipulations could not reliably be
separated from temporal clustering alone (reliable comparisons are reported in
the figure).

@@

At the group level, there seemed to be almost no lingering impact of sorting
early lists on memory for later lists. To simplify the presentation, we report
these null results in aggregate across the three feature groupings. Relative to
memory performance on late feature-rich lists, participants' memory performance
in all six order manipulation conditions showed no reliable differences
(semantic: $t(125) = 0.487,~p = 0.627,~d = 0.087,~\mathrm{CI} = [-1.661,~
2.323]$; lexicographic: $t(125) = 0.878,~p = 0.382,~d = 0.156,~\mathrm{CI} =
[-1.226,~3.044]$; visual: $t(126) = 1.437,~p = 0.153,~d = 0.254,~\mathrm{CI} =
[-0.447,~3.519]$). Nor did we observe any reliable differences in temporal
clustering on late lists (relative to late feature-rich lists; semantic:
$t(125) = 0.157,~p = 0.875,~d = 0.028,~\mathrm{CI} = [-1.859,~1.974]$;
lexicographic: $t(125) = 0.998,~p = 0.320,~d = 0.177,~\mathrm{CI} = [-0.902,~
2.920]$; visual: $t(126) = 0.548,~p = 0.585,~d = 0.097,~\mathrm{CI} = [-1.450,~
2.365]$). Aside from a slightly increased tendency for participants to cluster
words by their length on late visual order manipulation lists (more than late
feature-rich lists; $t(126) = 2.005,~p = 0.047,~d = 0.355,~\mathrm{CI} =
[0.211,~3.722]$), we observed no reliable differences in any type of feature
clustering on late order manipulation condition lists versus late feature-rich
lists ($|t|$s $\leq 1.124,~p$s $\geq 0.263$).

>>>>

%! suppress = LineBreak
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{geometry}
\usepackage[sort&compress]{natbib}
\usepackage{pxfonts}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{lineno}
\usepackage{amsopn}
\usepackage{rotating}
\usepackage{booktabs}

@@

% figures
\newcommand{\dynamicsRandom}{S1}
\newcommand{\dynamicsAdaptive}{S2}
\newcommand{\accuracyByList}{S3}
\newcommand{\clusterCorrs}{S4}
\newcommand{\fingerprintsRandom}{S5}
\newcommand{\fingerprintsAdaptive}{S6}
\newcommand{\recallInit}{S7}
\newcommand{\fingerprintTrajectoryRandom}{S8}
\newcommand{\featureCorrs}{S9}

% tables
\newcommand{\abbreviations}{S1}

@@

\begin{table}[tp]
\centering
\tiny
\input{tables/feature_rich_vs_reduced.tex}

\caption{\textbf{Comparing memory in the feature-rich versus reduced conditions
(all lists).} The $t$-tests reported in the table were carried out
across-participants, and reflect data aggregated across all lists from each
participant. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:feature-rich-vs-reduced}
\end{table}

To obtain a clean preliminary estimate of the consequences on memory of
randomly varying the font colors and locations of presented words (versus
holding the font color fixed at black, and the words' locations fixed at the
center of the screen), we compared participants' performance on the
\textit{feature-rich} and \textit{reduced} experimental conditions (see
\textit{Random conditions}, Fig.~\dynamicsRandom,
Tab.~\ref{tab:feature-rich-vs-reduced}). In the feature-rich condition, the
words' colors and locations varied randomly, and in the reduced condition,
words were always presented in black, at the center of the display. Aggregating
across all lists for each participant, we found no difference in recall
accuracy (i.e., the proportions of words successfully recalled) for
feature-rich versus reduced lists. However, participants in the feature-rich
condition clustered their recalls substantially more along every dimension we
examined (see \textit{Permutation-corrected feature clustering scores} for more
information about how we quantified each participant's clustering tendencies.)
Taken together, these comparisons suggest that adding new features changes how
participants organize their memories of studied words, even when those new
features are independent of the words themselves and vary randomly across
words. We found no evidence that those additional uninformative features were
distracting (in terms of their impact on memory performance), but they did
affect participants' recall dynamics (measured via their clustering scores).

@@

\begin{table}[tp]
\centering
\tiny
\input{tables/early_vs_late_fr.tex}

\caption{\textbf{Comparing memory for early versus late lists in the
feature-rich condition.} The $t$-tests reported in the table were carried out
across-participants. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:early-vs-late-feature-rich} 
\end{table}

\begin{table}[tp]
\centering
\tiny
\input{tables/early_vs_late_r.tex}

\caption{\textbf{Comparing memory for early versus late lists in the
reduced condition.} The $t$-tests reported in the table were carried out
across-participants. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:early-vs-late-reduced}
\end{table}

\begin{table}[tp]
\centering
\tiny
\input{tables/early_vs_late_early.tex}

\caption{\textbf{Comparing memory for early versus late lists in the
reduced (early) condition.} The $t$-tests reported in the table were carried out
across-participants. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:early-vs-late-early}
\end{table}

\begin{table}[tp]
\centering
\tiny
\input{tables/early_vs_late_late.tex}

\caption{\textbf{Comparing memory for early versus late lists in the
reduced (late) condition.} The $t$-tests reported in the table were carried out
across-participants. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:early-vs-late-late}
\end{table}

We next asked whether adding these incidental visual features to later lists
(after the participants had already studied impoverished lists), or removing
the visual features from later lists (after the participants had already
studied visually diverse lists) might affect memory performance. In other
words, we sought to test for potential effects of changing the ``richness'' of
participants' experiences over time. All participants studied and recalled a
total of 16 lists; we defined \textit{early} lists as the first eight lists and
\textit{late} lists as the last eight lists each participant encountered. To
help interpret our results, we compared participants' memories on early versus
late lists in the above feature-rich
(Tab.~\ref{tab:early-vs-late-feature-rich}) and reduced
(Tab.~\ref{tab:early-vs-late-reduced}) conditions. Participants in both
conditions remembered more words on early versus late lists. Participants in
the feature-rich (but not reduced) conditions exhibited more temporal
clustering on early versus late lists. And participants in both conditions
tended to exhibit more semantic clustering on early versus late lists.
Participants in the reduced (but not feature-rich) conditions tended to exhibit
more lexicographic clustering on early versus late lists. Taken together, these
comparisons suggest that even when the presence or absence of incidental visual
features is stable across lists, participants still exhibit some differences in
their performance and memory organization tendencies for early versus late
lists.

With these differences in mind, we next compared participants' memories on
early versus late lists for two additional experimental conditions (see
\textit{Random conditions}, Fig.~\dynamicsRandom). In a \textit{reduced
(early)} condition, we held the visual features constant on early lists, but
allowed them to vary randomly on late lists. In a \textit{reduced (late)}
condition, we allowed the visual features to vary randomly on early lists, but
held them constant on late lists. Given our above findings that (a)
participants tended to exhibit stronger clustering effects on feature-rich
(versus reduced) lists, and (b) participants tended to remember more words and
exhibit stronger clustering effects on early (versus late) lists, we expected
these early versus late differences to be enhanced in the reduced (early)
condition and diminished in the reduced (late) condition. However, to our
surprise, participants in \textit{neither} condition exhibited reliable
early-versus-late differences in accuracy, temporal clustering, nor
feature-based clustering
(Tabs.~\ref{tab:early-vs-late-early},~\ref{tab:early-vs-late-late}). We
hypothesized that adding or removing the variability in the visual features was
acting as a sort of ``event boundary'' between early and late
lists~\citep[e.g.,][]{ClewEtal19,RadvZack17,RadvCope06}. In prior work, we (and
others) have found that memories formed just after event boundaries can be
enhanced~\citep[e.g., due to less contextual interference between pre- and
post-boundary items;][]{MannEtal16, PettEtal16, GoldEtal17, FlorEtal17}.

\begin{table}[tp]
\centering
\tiny
\input{tables/feature_rich_vs_reduced_early.tex}

\caption{\textbf{Comparing memory in the feature-rich versus reduced (early)
conditions (all lists).} The $t$-tests reported in the table were carried out
across-participants, and reflect data aggregated across all lists from each
participant. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:feature-rich-vs-reduced-early}
\end{table}

\begin{table}[tp]
\centering
\tiny
\input{tables/reduced_vs_reduced_early.tex}

\caption{\textbf{Comparing memory in the reduced versus reduced (early)
conditions (all lists).} The $t$-tests reported in the table were carried out
across-participants, and reflect data aggregated across all lists from each
participant. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:reduced-vs-reduced-early}
\end{table}

\begin{table}[tp]
\centering
\tiny
\input{tables/feature_rich_vs_reduced_late.tex}

\caption{\textbf{Comparing memory in the feature-rich versus reduced (late)
conditions (all lists).} The $t$-tests reported in the table were carried out
across-participants, and reflect data aggregated across all lists from each
participant. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:feature-rich-vs-reduced-late}
\end{table}

\begin{table}[tp]
\centering
\tiny
\input{tables/reduced_vs_reduced_late.tex}

\caption{\textbf{Comparing memory in the reduced versus reduced (late)
conditions (all lists).} The $t$-tests reported in the table were carried out
across-participants, and reflect data aggregated across all lists from each
participant. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:reduced-vs-reduced-late}
\end{table}

We found that \textit{adding} incidental visual features on later lists that
had not been present on early lists (as in the reduced (early) condition)
served to enhance recall performance relative to conditions where all lists had
the same blends of features
(Tabs.~\ref{tab:feature-rich-vs-reduced-early},~\ref{tab:reduced-vs-reduced-early};
also see Fig.~\accuracyByList A). However, \textit{subtracting} irrelevant
visual features on later lists that \textit{had} been present on early lists
(as in the reduced (late) condition) did not appear to impact recall
performance
(Tabs.~\ref{tab:feature-rich-vs-reduced-late},~\ref{tab:reduced-vs-reduced-late})
These comparisons suggest that recall accuracy has a directional component:
accuracy is affected differently by removing features that had initially been
present versus adding features that had initially been absent. In contrast, we
found that participants exhibited more temporal and feature-based clustering
when we added incidental visual features to \textit{any} lists (feature-rich
versus reduced: Tab.~\ref{tab:feature-rich-vs-reduced}; reduced versus reduced
(early): Tab.~\ref{tab:reduced-vs-reduced-early}; reduced versus reduced
(late): Tab.~\ref{tab:reduced-vs-reduced-late}). Temporal and feature-based
clustering were not reliably different in the feature-rich versus reduced
(early) or reduced (late) conditions
(Tabs.~\ref{tab:feature-rich-vs-reduced-early},~\ref{tab:feature-rich-vs-reduced-late}).

@@

\begin{table}[tp]
\centering
\tiny
\input{tables/order_manip_vs_feature_rich_accuracy_early.tex}

\caption{\textbf{Comparing accuracy on early lists in the order manipulation
versus feature-rich conditions.} The $t$-tests reported in the table were
carried out across-participants, and reflect data aggregated across all early lists
from each participant. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:order-manip-vs-feature-rich-accuracy-early}
\end{table}

\begin{table}[tp]
\centering
\tiny
\input{tables/order_manip_vs_feature_rich_temporal_early.tex}

\caption{\textbf{Comparing temporal clustering on early lists in the order manipulation
versus feature-rich conditions.} The $t$-tests reported in the table were
carried out across-participants, and reflect data aggregated across all early lists
from each participant. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:order-manip-vs-feature-rich-temporal-early}
\end{table}

\begin{table}[tp]
\centering
\tiny
\input{tables/semantic_vs_feature_rich_feature_early.tex}

\caption{\textbf{Comparing feature-based clustering on early lists in the semantic order manipulation
versus feature-rich conditions.} The $t$-tests reported in the table were
carried out across-participants, and reflect data aggregated across all early lists
from each participant. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:semantic-vs-feature-rich-feature-early}
\end{table}

\begin{table}[tp]
\centering
\tiny
\input{tables/lexicographic_vs_feature_rich_feature_early.tex}

\caption{\textbf{Comparing feature-based clustering on early lists in the lexicographic order manipulation
versus feature-rich conditions.} The $t$-tests reported in the table were
carried out across-participants, and reflect data aggregated across all early lists
from each participant. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:lexicographic-vs-feature-rich-feature-early}
\end{table}

\begin{table}[tp]
\centering
\tiny
\input{tables/visual_vs_feature_rich_feature_early.tex}

\caption{\textbf{Comparing feature-based clustering on early lists in the visual order manipulation
versus feature-rich conditions.} The $t$-tests reported in the table were
carried out across-participants, and reflect data aggregated across all early lists
from each participant. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:visual-vs-feature-rich-feature-early}
\end{table}

@@

letter, color, and location). When we compared participants' memories for early
lists in each of these conditions to their memories for early lists in the
feature-rich condition
(Tab.~\ref{tab:order-manip-vs-feature-rich-accuracy-early}), we found that
participants in the category-ordered condition remembered more words than
participants in the feature-rich condition. Participants in the Participants in
the color-ordered condition also showed a trending increase in memory
performance on early lists. Participants' performances on early lists in all of
the other order manipulation conditions were indistinguishable from performance
on the early feature-rich lists. We also compared participants' temporal
clustering on early lists in each of these conditions to their temporal
clustering on early lists in the feature-rich condition
(Tab.~\ref{tab:order-manip-vs-feature-rich-temporal-early}). Participants in
both of the semantically ordered conditions exhibited stronger temporal
clustering on early lists (vs.~early feature-rich lists). Participants in the
length-ordered condition tended to exhibit \textit{less} temporal clustering on
early lists relative to early feature-rich lists, whereas participants in the
first letter-ordered condition exhibited stronger temporal clustering on early
lists. Participants in the visually ordered conditions showed similar temporal
clustering on early lists relative to early feature-rich lists. We also also
compared feature-based clustering on early lists across the order manipulation
and feature-rich conditions. Since these results were similar across both
semantic conditions (category and size;
Tab.~\ref{tab:semantic-vs-feature-rich-feature-early}), both lexicographic
conditions (length and first letter;
Tab.~\ref{tab:lexicographic-vs-feature-rich-feature-early}), and both visual
conditions (color and location;
Tab.~\ref{tab:visual-vs-feature-rich-feature-early}), here we aggregate data
from conditions that manipulated each of these three feature groupings in our
comparisons, to simplify the presentation. On early lists, participants in the
semantically ordered conditions exhibited stronger semantic clustering relative
to participants in the feature-rich condition, but showed no reliable
differences in lexicographic or visual clustering. Similarly, participants in
the lexicographically ordered conditions exhibited stronger (relative to
feature rich participants) lexicographic clustering on early lists, but showed
no reliable differences in semantic or visual clustering. And participants in
the visually ordered conditions exhibited stronger visual clustering (again,
relative to feature-rich participants, and on early lists), but showed no
reliable differences in semantic or lexicographic clustering. Taken together,
these order manipulation results suggest several broad patterns
(Figs.~\ref{fig:recall-dynamics}A,~\ref{fig:fingerprints}). First, most of the
order manipulations we carried out did \textit{not} reliably affect overall
recall performance. Second, most of the order manipulations increased
participants' tendencies to temporally cluster their recalls. Third, all of the
order manipulations enhanced participants' clustering of each condition's
target feature (i.e., semantic manipulations enhanced semantic clustering,
lexicographic manipulations enhanced lexicographic clustering, and visual
manipulations enhanced visual clustering; Fig.~\ref{fig:barplots}C) while
leaving clustering along other feature dimensions roughly unchanged (i.e.,
semantic manipulations did not affect lexicographic or visual clustering, and
so on). Although it is not possible to fully separate feature-based versus
temporal clustering when considering sorted lists, we used a permutation-based
procedure to identify the degree of feature clustering over and above what
could be accounted for by temporal clustering alone (see \textit{Factoring out
the effects of temporal clustering}). When we carried out this analysis
(Fig.~\ref{fig:barplots}D), we found that participants exhibited more semantic
clustering on semantically sorted lists than on randomly ordered lists, but the
effects of the other order manipulations could not reliably be separated from
temporal clustering alone (reliable comparisons are reported in the figure).

@@

\begin{table}[tp]
\centering
\tiny
\input{tables/carryover_accuracy.tex}

\caption{\textbf{Comparing accuracy on late lists in order-manipulatation
versus feature-rich conditions.} The $t$-tests reported in the table were
carried out across-participants, and reflect data aggregated across all late
lists from each participant. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:accuracy-carryover}
\end{table}

\begin{table}[tp]
\centering
\tiny
\input{tables/carryover_temporal_clustering.tex}

\caption{\textbf{Comparing temporal clustering on late lists in order-manipulatation
versus feature-rich conditions.} The $t$-tests reported in the table were
carried out across-participants, and reflect data aggregated across all late
lists from each participant. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:temporal-carryover}
\end{table}

\begin{table}[tp]
\centering
\tiny
\input{tables/semantic_carryover_feature_clustering.tex}

\caption{\textbf{Comparing feature-based clustering on late lists in semantic order-manipulatation
versus feature-rich conditions.} The $t$-tests reported in the table were
carried out across-participants, and reflect data aggregated across all late
lists from each participant. Abbreviations used in this table are defined in
Table~\abbreviations.}

\label{tab:semantic-feature-carryover}
\end{table}

\begin{table}[tp]
\centering
\tiny
\input{tables/lexicographic_carryover_feature_clustering.tex}

\caption{\textbf{Comparing feature-based clustering on late lists in
lexicographic order-manipulatation versus feature-rich conditions.} The
$t$-tests reported in the table were carried out across-participants, and
reflect data aggregated across all late lists from each participant.
Abbreviations used in this table are defined in Table~\abbreviations.}

\label{tab:lexicographic-feature-carryover}
\end{table}
